Filling in data gaps on refugee children: Statistical imputation for missing count data

### 2023-01-16 paper.qmd
update tables and figures, add group level tables to annex

imputations <- imputations %>% 
   mutate(female = rowSums(select(., female_0_4:female_60)),
          male = rowSums(select(., male_0_4:male_60)),
          children = rowSums(select(., female_0_4:female_12_17, male_0_4:male_12_17)),
          adults = rowSums(select(., female_18_59, female_60, male_18_59, male_60)),
          childrenProp = children/total,
          femaleProp = female/total,
          origin_countryShort = case_when(
              origin_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              origin_country == "United Kingdom of Great Britain and Northern Ireland" ~ "Bolivia",
              origin_country == "Bolivia (Plurinational State of)" ~ "UK",
              origin_country == "Central African Republic" ~ "CAR",
              origin_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              origin_country == "Lao People's Democratic Republic" ~ "Lao",
              origin_country == "Russian Federation" ~ "Russia",
              origin_country == "Syrian Arab Republic" ~ "Syria",
              origin_country == "Democratic Republic of the Congo" ~ "DRC",
              origin_country == "United Republic of Tanzania" ~ "Tanzania",
              origin_country == "United States of America" ~ "USA",
              origin_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(origin_country)
          ),
         asylum_countryShort = case_when(
              asylum_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              asylum_country == "United Kingdom of Great Britain and Northern Ireland" ~ "Bolivia",
              asylum_country == "Bolivia (Plurinational State of)" ~ "UK",
              asylum_country == "Central African Republic" ~ "CAR",
              asylum_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              asylum_country == "Lao People's Democratic Republic" ~ "Lao",
              asylum_country == "Russian Federation" ~ "Russia",
              asylum_country == "Syrian Arab Republic" ~ "Syria",
              asylum_country == "Democratic Republic of the Congo" ~ "DRC",
              asylum_country == "United Republic of Tanzania" ~ "Tanzania",
              asylum_country == "United States of America" ~ "USA",
              asylum_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(asylum_country)  
          )
         ) %>% 
  mutate(origin_countryShort = factor(origin_countryShort), 
         asylum_countryShort = factor(asylum_countryShort))

@@

```{r, message=FALSE, warning=FALSE, out.width="80%", out.height="180%"}
#| label: fig-missingAsy
#| fig-cap: Missingness by region of asylum
#| warning: false
#| error: false
#| echo: false
#| message: false

@@

t.worldregion.children %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 1) %>%
  tab_options(table.font.size = "40%") %>%
  cols_label(asylum_region = md("Region<br>of asylum"),
         childrenProp=  md("Percentage of<br>children in<br>available data"),
        childrenMeanProp =md("Estimated<br>percentage with<br>imputed data" ) ,
       lowerUI  =md("Lower 95%<br>uncertainty") ,
        upperUI =md("Upper 95%<br>uncertainty"))

@@

p.imputations.syr.children <- ggplot(data = imputations.syr %>% filter(asylum_iso3 %in% c("SWE", "USA",    "GBR", "GRC")),
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_slab() +
  labs(x = "Proportion of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank())

@@

p.imputations.ven.children <- ggplot(data = imputations.ven %>% filter(asylum_iso3 %in% c("ARG", "ECU", "CHL", "BRA"), popType=="VDA"),
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_slab() +
  labs(x = "Proportion of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank())

@@

```{r, fig.pos = "H"}
#| label: fig-impChildSyrReg
#| fig-cap: Uncertainty in posterior predictive distribution of the proportion of children among Syrian refugees across regions
#| fig-height: 13
#| fig-width: 9
#| warning: false
#| error: false
#| echo: false
#| message: false
p.imputations.syr.reg.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_pointinterval() +
  labs(x = "Proportion of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 13, angle = 0),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank()) +
    facet_wrap(~ asylum_region, scales = "free", ncol = 2) +
  theme(strip.text = element_text(size = 16))

@@

# Annex I: Model output
```{r}
#| label: tbl-model_fixed
#| tbl-cap: Model - fixed effects
#| warning: false
#| error: false
#| echo: false
#| message: false
m.ageonly.sum <- summary(m.ageonly)
m.ageonly.ranefsum <- t(m.ageonly.sum$random %>% as_tibble() )
colnames(m.ageonly.ranefsum) <- c("female_0_4", "female_5_11" , "female_12_17" , "female_60" ,
                                  "male_0_4"  ,               "male_5_11"   ,            
"male_12_17"     ,          "male_18_59"     ,          "male_60")

m.ageonly.ranefsum <- m.ageonly.ranefsum %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(`Group and summary`=rowname)

m.ageonly.fixedsum <- m.ageonly.sum$fixed

m.ageonly.fixedsum %>% gt()

@@

>>>>

imputations <- imputations %>% 
   mutate(female = rowSums(select(., female_0_4:female_60)),
          male = rowSums(select(., male_0_4:male_60)),
          children = rowSums(select(., female_0_4:female_12_17, male_0_4:male_12_17)),
          adults = rowSums(select(., female_18_59, female_60, male_18_59, male_60)),
          childrenProp = children/total,
          femaleProp = female/total,
          origin_countryShort = case_when(
              origin_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              origin_country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
              origin_country == "Bolivia (Plurinational State of)" ~ "Bolivia",
              origin_country == "Central African Republic" ~ "CAR",
              origin_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              origin_country == "Lao People's Democratic Republic" ~ "Lao",
              origin_country == "Russian Federation" ~ "Russia",
              origin_country == "Syrian Arab Republic" ~ "Syria",
              origin_country == "Democratic Republic of the Congo" ~ "DRC",
              origin_country == "United Republic of Tanzania" ~ "Tanzania",
              origin_country == "United States of America" ~ "USA",
              origin_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(origin_country)
          ),
         asylum_countryShort = case_when(
              asylum_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              asylum_country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
              asylum_country == "Bolivia (Plurinational State of)" ~ "Bolivia",
              asylum_country == "Central African Republic" ~ "CAR",
              asylum_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              asylum_country == "Lao People's Democratic Republic" ~ "Lao",
              asylum_country == "Russian Federation" ~ "Russia",
              asylum_country == "Syrian Arab Republic" ~ "Syria",
              asylum_country == "Democratic Republic of the Congo" ~ "DRC",
              asylum_country == "United Republic of Tanzania" ~ "Tanzania",
              asylum_country == "United States of America" ~ "USA",
              asylum_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(asylum_country)  
          )
         ) %>% 
  mutate(origin_countryShort = factor(origin_countryShort), 
         asylum_countryShort = factor(asylum_countryShort))

@@

```{r, message=FALSE, warning=FALSE, out.width="80%", out.height="180%"}
#| label: fig-missingAsy
#| fig-cap: Missingness by region of asylum
#| fig-height: 6
#| fig-width: 6
#| warning: false
#| error: false
#| echo: false
#| message: false

@@

t.worldregion.children %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 1) %>%
  tab_options(table.font.size = "40%") %>%
  cols_label(asylum_region = md("Region"),
         childrenProp=  md("% available data"),
        childrenMeanProp =md("% estimated" ) ,
       lowerUI  =md("Lower 95% CI") ,
        upperUI =md("Upper 95% CI")) %>% 
    tab_style(
    style = cell_borders(
      sides = c("right"),
      color = "	#c1cdcd",
      weight = 0.7,
      style = "dashed"
    ),
    locations = cells_body(
      columns = 2,
      rows = everything()
    )
  ) %>% 
    tab_style(
    style = cell_borders(
      sides = c("right"),
      color = "	#c1cdcd",
      weight = 0.7,
      style = "dashed"
    ),
locations = cells_column_labels(
      columns = 2
    )
  )

@@

p.imputations.syr.children <- ggplot(data = imputations.syr %>% filter(asylum_iso3 %in% c("SWE", "USA",    "GBR", "GRC")),
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_slab() +
  labs(x = "Percentage of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank())

@@

p.imputations.ven.children <- ggplot(data = imputations.ven %>% filter(asylum_iso3 %in% c("ARG", "ECU", "CHL", "BRA"), popType=="VDA"),
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_slab() +
  labs(x = "Percentage of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank())

@@

```{r, fig.pos = "H"}
#| label: fig-impChildSyrReg
#| fig-cap: Uncertainty in posterior predictive distribution of the proportion of children among Syrian refugees across regions
#| fig-height: 13
#| fig-width: 9
#| warning: false
#| error: false
#| echo: false
#| message: false
p.imputations.syr.reg.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_pointinterval() +
  labs(x = "Percentage of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 13, angle = 0),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank()) +
    facet_wrap(~ asylum_region, scales = "free", ncol = 2) +
  theme(strip.text = element_text(size = 16))

@@

# Annex I: Model output
```{r}
#| label: tbl-model_fixed
#| tbl-cap: Model - fixed effects
#| warning: false
#| error: false
#| echo: false
#| message: false
m.ageonly.sum <- summary(m.ageonly)

m.ageonly.fixedsum <- m.ageonly.sum$fixed  %>% 
  rownames_to_column() %>% 
  select(rowname, Estimate, Est.Error, `l-95% CI`,   `u-95% CI`) %>%
  rename(Coef = rowname)

m.ageonly.fixedsum %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 2)

@@

```{r}
#| label: tbl-model_re1
#| tbl-cap: Model - origin country group level standard deviation 
#| warning: false
#| error: false
#| echo: false
#| message: false
 m.ageonly.sum$random [1]$origin_iso3 %>% 
  rownames_to_column() %>% 
  as_tibble() %>% 
  select(rowname, Estimate, Est.Error, `l-95% CI`,   `u-95% CI`) %>%
  rename(Coef = rowname) %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 2)

```{r}
#| label: tbl-model_re2
#| tbl-cap: Model - asylum region group level standard deviation 
#| warning: false
#| error: false
#| echo: false
#| message: false
 m.ageonly.sum$random [2]$`origin_iso3:asylum_sdgregion` %>% 
  rownames_to_column() %>% 
  as_tibble() %>% 
  select(rowname, Estimate, Est.Error, `l-95% CI`,   `u-95% CI`) %>%
  rename(Coef = rowname) %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 2)

```{r}
#| label: tbl-model_re3
#| tbl-cap: Model - asylum country group level standard deviation 
#| warning: false
#| error: false
#| echo: false
#| message: false
 m.ageonly.sum$random [3]$`origin_iso3:asylum_sdgregion:asylum_iso3` %>% 
  rownames_to_column() %>% 
  as_tibble() %>% 
  select(rowname, Estimate, Est.Error, `l-95% CI`,   `u-95% CI`) %>%
  rename(Coef = rowname) %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 2)
